MSstatsPTM: Statistical relative quantification of post-translational modifications in bottom-up mass spectrometry-based proteomics

### 2022-06-30 main/MSstatsPTM_Main.tex
Added refs

\medskip \noindent The two versions of $t$-test \todo{OV: ANOVA?} above can be expanded with Empirical Bayes moderation in Limma~\cite{ Ritchie_15a, Schwammle2015, Smyth:2005, Zhu,Chappell:2021}. In particular, the ratio-based version models the ratios $\hat{u}_{ij}$ in~\eqref{adj_estimation}. Limma then fits a linear model using the transformed inputs.

@@

$MSstats$~\cite{Choi:2014} is designed for protein significance analysis and statistical relative quantification of proteins and peptides in global, targeted and data-independent proteomics. The package takes as input a list of log$_2$-intensities $y_{ijk}$, used to characterize the identified protein. For each protein, the feature log-intensities are first summarized into a single value per protein per run in the subplot model. This is done using Tukey's median polish (TMP)~\cite{Tukey:1977}, a simple and robust procedure, which iteratively fits a two-way additive model with the effects of run and feature, summarizing them into a single value per run. These summarized values are then used as input to fit a linear model based on the family of linear mixed-effects models \cite{Bolker2009} \cite{Faraway:2006}. For example, the model for a TMT experiment would be as follows

>>>>

\medskip \noindent The two versions of $t$-test \todo{OV: ANOVA?} above can be expanded with Empirical Bayes moderation in Limma~\cite{ Ritchie_15a, Schwammle2015, Smyth:2004, Smyth:2005, Zhu,Chappell:2021}. In particular, the ratio-based version models the ratios $\hat{u}_{ij}$ in~\eqref{adj_estimation}. Limma then fits a linear model using the transformed inputs.

@@

$MSstats$~\cite{Choi:2014} is designed for protein significance analysis and statistical relative quantification of proteins and peptides in global, targeted and data-independent proteomics. The package takes as input a list of log$_2$-intensities $y_{ijk}$, used to characterize the identified protein. For each protein, the feature log-intensities are first summarized into a single value per protein per run in the subplot model. This is done using Tukey's median polish (TMP)~\cite{Tukey:1977}, a simple and robust procedure, which iteratively fits a two-way additive model with the effects of run and feature, summarizing them into a single value per run. These summarized values are then used as input to fit a linear model based on the family of linear mixed-effects models \cite{McLean:1991, Faraway:2006, Bolker2009}. For example, the model for a TMT experiment would be as follows
